---
title: "Resp Support at three year (choice=Decannulated) of Tracheostomy Resource Utilization Study"
output:
  html_document:
    toc: yes
    toc_depth: 3
---
```{r setup, echo=FALSE}
knitr::opts_chunk$set(eval = TRUE,
echo = FALSE,
cache = FALSE,
include = TRUE,
collapse = TRUE,
warning = FALSE,
message = FALSE,
dependson = NULL,
engine = "R", # Chunks will always have R code, unless noted
error = TRUE,
fig.path="Figures/nas_mwph", # Set the figure optins
fig.align = "center",
fig.width = 6,
fig.height = 6,
dpi = 300)
```

```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
library(table1) 
trach <- read.csv("C:\\Users\\15715\\Desktop\\22fall\\Mathangi\\trachstudy\\TracheostomyResource_DATA_LABELS_2023-02-03_1248.csv")
```

```{r}
my_theme <- function() {
 theme(axis.text.x=element_text(vjust=1, size=11, color="black"),
 axis.text.y=element_text(hjust=1,vjust=1, size=12, color="black"),
 panel.background=element_rect(fill="white",color="black"),
 plot.title = element_text(size = 15, face = "bold"),
 panel.grid.minor.y = element_blank(),
 panel.grid.major.y = element_line(color="#CCCCCC", linetype = 2),
 panel.grid.minor.x = element_blank(),
 panel.grid.major.x = element_line(color="#CCCCCC", linetype = 2),
 axis.title.x=element_text(size=22, face = "bold"),
 axis.title.y=element_text(size=22, face = "bold"),
 strip.text=element_text(size=12, face = "bold"),
 legend.text=element_text(size=20,face="bold"),
 legend.title=element_text(colour="black", size=22, face="bold"),
 axis.line.x = element_line(size = 1.5),
 axis.line.y = element_line(size = 1.5),
 axis.ticks = element_line(size = 1.5,color="black"),
 axis.text = element_text(color="black"),
 axis.ticks.length= unit(0.45, "cm"),
 legend.key.width = unit(1.2, "cm"))
 #legend.background = element_rect(fill="white",color="black"))
}
```

```{r}
trach1 <- data.frame(fssmental_year1=trach$Functional.Status.Score..MENTAL, fsssensory_year1=trach$Functional.Status.Score..SENSORY, fsscommunication_year1=trach$Functional.Satus.Score..COMMUNICATION, 
fssmotor_year1=trach$Functional.Status.Scale..MOTOR.,
fssfeeding_year1=trach$Functional.Status.Scale..FEEDING,
fssresp_year1=trach$Functional.Status.Scale..RESPIRATORY,
respsupports_Decan_year1=trach$Resp.Support.at.one.year..choice.Decannulated.,

fssmental_year3=trach$Functional.Status.Scale..MENTAL,
fsssensory_year3=trach$Functional.Status.Scale..SENSORY,
fsscommunication_year3=trach$Functional.Status.Scale..COMMUNICATION,
fssmotor_year3=trach$Functional.Status.Scale..MOTOR,
fssfeeding_year3=trach$Functional.Status.Scale..FEEDING.1,
fssresp_year3=trach$Functional.Status.Scale..RESPIRATORY.1,
respsupports_Decan_year3=trach$Resp.Support.at.three.years..choice.Decannulated.)
```

## Bar plots of 6 components of FSS total
1: Normal; 2: Mild; 3: Moderate; 4: Severe; 5: Very severe dysfunction\
13 missing in the 6 components
```{r}
# resp <- data.frame(mental=trach$Functional.Status.Scale..MENTAL,
# sensory=trach$Functional.Status.Scale..SENSORY,
# communication=trach$Functional.Status.Scale..COMMUNICATION,
# motor=trach$Functional.Status.Scale..MOTOR,
# feeding=trach$Functional.Status.Scale..FEEDING.1,
# resp=trach$Functional.Status.Scale..RESPIRATORY.1,
# respsupports_Decan=trach$Resp.Support.at.three.years..choice.Decannulated.)
resp <- data.frame(mental=trach$Functional.Status.Score..MENTAL,
sensory=trach$Functional.Status.Score..SENSORY,
communication=trach$Functional.Satus.Score..COMMUNICATION,
motor=trach$Functional.Status.Scale..MOTOR.,
feeding=trach$Functional.Status.Scale..FEEDING,
resp=trach$Functional.Status.Scale..RESPIRATORY,
respsupports_Decan=trach$Resp.Support.at.three.years..choice.Decannulated.)

num <- function(x) {
  ifelse(grepl("^Normal", x), 1, 
         ifelse(grepl("^Mild", x), 2, 
                ifelse(grepl("^Moderate", x), 3, 
                       ifelse(grepl("^Severe", x), 4, 
                              ifelse(grepl("^Very", x), 5, NA)))))
}

resp1 <- data.frame(cbind(apply(resp[,-c(7)], 2, function(x) num(x))), respsupports_Decan=resp[,7])
resp1[,c(1:6)] <- lapply(resp1[,c(1:6)], as.factor)
resp1[,c(7)] <- as.factor(ifelse(resp1[,c(7)]=="Checked",1,0))

resp2 <- na.omit(resp1)

df_long <- resp2[,c(1:6)] %>%
  gather(key = "components", value = "value")

df_long <- cbind(id=factor(rev(paste0("#", 1:50)), level = rev(paste0("#", 1:50))),df_long)
# Create bar plots using facet_wrap
ggplot(df_long, aes(x = value, fill = components)) +
  geom_bar(stat = "count", position = "dodge") +
  ylim(c(0,50)) +
  facet_wrap(~ components, ncol = 3) + 
  theme(legend.position="none") +
  # geom_text(aes(label = paste("Count:", (df_long %>% count(value, variable))$n)), position = position_dodge(width = 0.9), vjust = -0.5)+
  # geom_text(data = df_long %>% count(value, components), aes(label = paste("Count:",n), y = Inf, x  = -Inf), position = position_dodge(width = 0.9), vjust = -0.5)+
  geom_text(stat = "count", aes(label = ..count..), vjust = 0) +
  labs(x = "Condition", y = "Count", title = "Facet plots of 6 components of FSS total in year 3") +
  my_theme()
```

## Heatmap-like plot of 6 components 
```{r}
ggplot(df_long, aes(components, id)) + geom_tile(aes(fill = value)) + scale_fill_discrete(name = "Condition", labels = c("Normal", "Mild", "Moderate", "Severe", "Very severe"))

# ggsave("heat.png", path = "C:\\Users\\15715\\Desktop\\22fall\\Mathangi\\trachstudy", width = 20, height = 20, device='tiff', dpi=20)
```


## Fit logistic regression for all 6 components (outcome is Resp Support at tree year (choice=Decannulated))
All covariates are insignificant. 
```{r warning=FALSE}
convert <- function(x) ifelse(x==1,1,"other")
resp3 <- cbind(as.data.frame(lapply(resp2[,1:6], convert)), respsupports_Decan=resp2$respsupports_Decan)

my_factor <- function(x) factor(x, levels = c("other","1"))
resp3[,1:6] <- lapply(resp3[1:6], my_factor)

model1 <- glm(respsupports_Decan ~ mental + sensory + communication + motor + feeding + resp, family = "binomial", data=resp3)
summary(model1)
```

## Fit logistic regression for one component(MOTOR, FEEDING)
Going to Normal from other categories in MOTOR, the odds of checked in the outcome (versus unchecked) increase by a factor of 4.4.\

Going to Normal from other categories in FEEDING, the odds of checked in the outcome (versus unchecked) increase by a factor of 7.5.
```{r warning=FALSE}
model2 <- glm(respsupports_Decan ~ motor, family = "binomial", data=resp3)
summary(model2)

model3 <- glm(respsupports_Decan ~ feeding, family = "binomial", data=resp3)
summary(model3)

# model3 <- glm(respsupports_Decan ~ resp, family = "binomial", data=resp3)
# summary(model3)

table(resp3$respsupports_Decan)

confint.default(model2)
confint.default(model3)

summary(glm(respsupports_Decan ~ mental, family = "binomial", data=resp3))
summary(glm(respsupports_Decan ~ sensory, family = "binomial", data=resp3))
summary(glm(respsupports_Decan ~ communication, family = "binomial", data=resp3))
summary(glm(respsupports_Decan ~ resp, family = "binomial", data=resp3))
```
## Stepwise of logistic regression in terms of AIC
step1: mental + sensory + communication + motor + feeding + resp\
step2: mental + sensory + communication + motor + resp\
step3: mental + sensory + communication + resp\
step4: mental + sensory + resp\
step5: sensory + resp\
```{r}
step(model1,direction = c("both"))
```
without resp
```{r}
model1.1 <- glm(respsupports_Decan ~ mental + sensory + communication + motor + feeding, family = "binomial", data=resp3)
step(model1.1,direction = c("both"))
```

```{r}
model1.2 <- glm(respsupports_Decan ~ sensory + feeding, family = "binomial", data=resp3)
summary(model1.2)
confint.default(model1.2)
```

```{r}
model4 <- glm(respsupports_Decan ~ motor+feeding, family = "binomial", data=resp3)
summary(model4)
```

When resp=="Normal", respsupports_Decan=="checked"
```{r}
# cor(as.numeric(resp3$resp), as.numeric(resp3$respsupports_Decan), method = "spearman")

resp2$resp
resp2$respsupports_Decan
```

## P-value of chi-square test between one component and the outcome
FEEDING and RESPIRATORY have relationship with the outcome
```{r warning=FALSE}
p <- function(x) {
  table_data <- table(x, resp3$respsupports_Decan)
return(chisq.test(table_data)$p.value)}
# Perform a chi-square test of independence
lapply(resp3[1:6], p)
```