---
title: "Descriptive Analysis of Tracheostomy Resource Utilization Study"
output:
  html_document:
    toc: yes
    toc_depth: 3
---
```{r setup, echo=FALSE}
knitr::opts_chunk$set(eval = TRUE,
echo = FALSE,
cache = FALSE,
include = TRUE,
collapse = TRUE,
warning = FALSE,
message = FALSE,
dependson = NULL,
engine = "R", # Chunks will always have R code, unless noted
error = TRUE,
fig.path="Figures/nas_mwph", # Set the figure optins
fig.align = "center",
fig.width = 6,
fig.height = 6,
dpi = 300)
```

```{r}
my_theme <- function() {
 theme(axis.text.x=element_text(vjust=1, size=11, color="black"),
 axis.text.y=element_text(hjust=1,vjust=1, size=12, color="black"),
 panel.background=element_rect(fill="white",color="black"),
 plot.title = element_text(size = 15, face = "bold"),
 panel.grid.minor.y = element_blank(),
 panel.grid.major.y = element_line(color="#CCCCCC", linetype = 2),
 panel.grid.minor.x = element_blank(),
 panel.grid.major.x = element_line(color="#CCCCCC", linetype = 2),
 axis.title.x=element_text(size=22, face = "bold"),
 axis.title.y=element_text(size=22, face = "bold"),
 strip.text=element_text(size=12, face = "bold"),
 legend.text=element_text(size=20,face="bold"),
 legend.title=element_text(colour="black", size=22, face="bold"),
 axis.line.x = element_line(size = 1.5),
 axis.line.y = element_line(size = 1.5),
 axis.ticks = element_line(size = 1.5,color="black"),
 axis.text = element_text(color="black"),
 axis.ticks.length= unit(0.45, "cm"),
 legend.key.width = unit(1.2, "cm"))
 #legend.background = element_rect(fill="white",color="black"))
}
```

```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
library(table1) 
library(ggpubr)
trach <- read.csv("C:\\Users\\15715\\Desktop\\22fall\\Mathangi\\trachstudy\\TracheostomyResource_DATA_LABELS_2023-02-03_1248.csv")
```
## Interested variables\

Covariates:\
demographics(record_id, age, race, insurance)\
Mortality before one-year follow-up\
Mortality before three-year follow-up\
Indication for tracheostomy radio\
Total hospital stay (days)\
Duration (in days) of continuous infusion of sedatives\
Duration (in days) of continuous infusion of vasoactives\
Pediatric Cerebral Performance Category(Admission, Discharge, year1, year3, year5, mortality)\
Functional Status Score TOTAL(pre-Admission, admission, discharge, year1, year3, year5, mortality)\

Outcomes:\
Number of hospital admissions(year1, year3, year5)\
Number of ED visits(year1, year3, year5)\
Number of Prescription Medications(Discharge, year1, year3, year5)\
```{r}
trach1 <- data.frame(record_id=trach$Record.ID, ageattrach=trach$Age.at.tracheostomy, age_year1=trach$Age.at.One.year.Follow.up..years., 
age_year3=trach$Age.at.Three.year.Follow.up,
age_year5=trach$Age.at.Five.year.Follow.up,
age_death=trach$Age.at.Death..in.months.,
mortality1=trach$Mortality.before.one.year.follow.up.,
mortality3=trach$Mortality.prior.to.3.year.follow.up.,
mortality5=trach$Mortality.before.5.year.follow.up.,
race=trach$Race, insurance=trach$Type.of.Insurance,


indication_fortrach=trach$Indication.for.tracheostomy, hospital_lengthofstay=trach$Total.hospital.stay..days., sedatives=trach$Duration..in.days..of.continuous.infusion.of.sedatives, vasoactives=trach$Duration..in.days..of.continuous.infusion.of.vasoactives,

admit_pcpc=trach$Admission.Pediatric.Cerebral.Performance.Category,
discharge_pcpc=trach$Discharge.Pediatric.Cerebral.Performance.Category,
pcpc_year1=trach$Pediatric.Cerebral.Performance.Category.1.year.post.tracheostomy, pcpc_year3=trach$Pediatric.Cerebral.Performance.Category.at.3.years.post.tracheostomy, pcpc_year5=trach$Pediatric.Performance.Category.at.5.years.post.tracheostomy,
pcpc_mortality=trach$Pediatric.Cerebral.Performance.Category,

admitfss_totalscore=trach$Admission.Functional.Status.Score.TOTAL,
dischargefss_total=trach$Discharge.Functional.Status.Score.TOTAL,
fsstotal_year1=trach$Functional.Status.Scale.TOTAL.SCORE,
fsstotal_year3=trach$Functional.Status.Scale..TOTAL.,
fsstotal_year5=trach$Functional.Status.Scale..TOTAL.score,
fsstotal_mortality=trach$Functional.Status.Scale..TOTAL.SCORE..,

admit_year1=trach$Number.of.hospital.admissions, admit_year3=trach$Number.of.hospital.admissions.1, admit_year5=trach$Number.of.hospital.admissions.2, 
edvisits_year1=trach$Number.of.ED.visits, 
edvisits_year3=trach$Number.of.ED.visits.1, 
edvisits_year5=trach$Number.of.ED.visits.2,
medsatdischarge=trach$Total.number.of.prescription.medications.at.discharge,
meds_year1=trach$Number.of.Prescription.Medications, meds_year3=trach$Number.of.Prescription.Medications.1, meds_year5=trach$Number.of.Prescription.Medications.2)

for (i in 1:nrow(trach1)) {
  trach1$admit_pcpc[i] <- substr(trach1$admit_pcpc[i], start = 1, stop = 1)
  trach1$discharge_pcpc[i] <- substr(trach1$discharge_pcpc[i], start = 1, stop = 1)
  trach1$pcpc_year1[i] <- substr(trach1$pcpc_year1[i], start = 1, stop = 1)
  trach1$pcpc_year3[i] <- substr(trach1$pcpc_year3[i], start = 1, stop = 1)
  trach1$pcpc_year5[i] <- substr(trach1$pcpc_year5[i], start = 1, stop = 1)
  trach1$pcpc_mortality[i] <- substr(trach1$pcpc_mortality[i], start = 1, stop = 1)
}
trach1[, 15:20] <- replace(trach1[, 15:20], trach1[, 15:20] == 1, "Normal")
trach1[, 15:20] <- replace(trach1[, 15:20], trach1[, 15:20] == 2, "Mild")
trach1[, 15:20] <- replace(trach1[, 15:20], trach1[, 15:20] == 3, "Moderate")
trach1[, 15:20] <- replace(trach1[, 15:20], trach1[, 15:20] == 4, "Severe")
trach1[, 15:20] <- replace(trach1[, 15:20], trach1[, 15:20] == 5, "Coma")
f <- function(x) factor(x, levels = c("Normal", "Mild", "Moderate", "Severe", "Coma"))
trach1[,15:20] <- lapply(trach1[,15:20], f) 
trach1[trach1 == ""] <- NA
trach1$ageattrach <- as.numeric(difftime(trach$Date.of.Tracheostomy, trach$Date.of.Birth, units = "days")/30)

trach1$age_year1 <- ifelse(trach1$mortality1=="Yes"|trach1$mortality1=="", NA, trach1$ageattrach+12)
trach1$age_year3 <- ifelse(trach1$mortality3=="Yes"|trach1$mortality3=="", NA, trach1$ageattrach+36)
trach1$mortality3 <- ifelse(trach1$mortality1=="Yes", "Yes", trach1$mortality3)
# 
# sum(trach1$ageattrach<3)
# sum(3<=trach1$ageattrach& trach1$ageattrach<8)
# sum(trach1$ageattrach>=8)
# 
# table(trach1$mortality1)
# table(trach1$mortality3)
# table(trach1$mortality5)
```

## Plots of categorical variables and outcomes

### Races
```{r}
ggplot(trach1, aes(x = race, fill= race)) +
  geom_bar(aes(y = after_stat(count / sum(count)*100))) +
  xlab("Race") +
  ylab("Percentage of Race")
```

### Indication for tracheostomy radio
```{r}
ggplot(trach1, aes(x = indication_fortrach, fill= indication_fortrach)) +
  geom_bar(aes(y = after_stat(count / sum(count)*100))) +
  scale_x_discrete(guide = guide_axis(n.dodge=3)) +
  xlab("Indication for tracheostomy radio") +
  ylab("Percentage of Indication for tracheostomy radio")
```

### Pediatric Cerebral Performance Category
```{r}
ggplot(trach1, aes(x = admit_pcpc, fill= admit_pcpc)) +
  geom_bar(aes(y = after_stat(count / sum(count)*100))) +
  xlab("Admission Pediatric Cerebral Performance Category") +
  ylab("Percentage of Admission Pediatric Cerebral Performance Category")

ggplot(trach1, aes(x = discharge_pcpc, fill= discharge_pcpc)) +
  geom_bar(aes(y = after_stat(count / sum(count)*100))) +
  xlab("Discharge Pediatric Cerebral Performance Category") +
  ylab("Percentage of Discharge Pediatric Cerebral Performance Category")

ggplot(trach1, aes(x = pcpc_year1, fill= pcpc_year1)) +
  geom_bar(aes(y = after_stat(count / sum(count)*100))) +
  xlab("Pediatric Cerebral Performance Category 1-year posttracheostomy") +
  ylab("Percentage of Pediatric Cerebral Performance Category 1-year post-tracheostomy")

ggplot(trach1, aes(x = pcpc_year3, fill= pcpc_year3)) +
  geom_bar(aes(y = after_stat(count / sum(count)*100))) +
  xlab("Pediatric Cerebral Performance Category at 3 years post-tracheostomy") +
  ylab("Percentage of Pediatric Cerebral Performance Category at 3 years post-tracheostomy")

ggplot(trach1, aes(x = pcpc_year5, fill= pcpc_year5)) +
  geom_bar(aes(y = after_stat(count / sum(count)*100))) +
  xlab("Pediatric Performance Category at 5 years post tracheostomy") +
  ylab("Percentage of Pediatric Performance Category at 5 years post-tracheostomy")

ggplot(trach1, aes(x = pcpc_mortality, fill= pcpc_mortality)) +
  geom_bar(aes(y = after_stat(count / sum(count)*100))) +
  xlab("Pediatric Cerebral Performance Category-mortality") +
  ylab("Percentage of Pediatric Cerebral Performance Category-mortality")
```

### Outcomes
```{r}
filtered <- trach1 %>%
  select(starts_with("admit"), starts_with("meds"), starts_with("edvisits")) %>% 
  select(-1, -2)
filtered <- filtered[, !sapply(filtered, is.factor)]
filtered <- filtered[, !sapply(filtered, is.character)]
means <- as.vector(colMeans(filtered, na.rm = TRUE))
# var <- substr(colnames(filtered), 1, nchar(colnames(filtered))-6)
var <- rep(c("hospital_admissions","Prescription_Medications","ED_visits"), each=4)
means <- c(NA, 3.0400000, 1.7948718, 1.1250000, 10.6981132, 8.3000000, 6.8421053, 6.0000000, NA, 1.4800000, 0.8717949, 0.5625000)
aim2 <- data.frame(Category=var, Mean=means, Year=rep(c("discharge", "year1","year3","year5"), 3))
# as.vector(apply(filtered, 2, function(x) sum(is.na(x))))
aim2$Missing <- c(53, 3, 14, 37,  0,  3, 15, 37, 53, 3, 14, 37)
ggplot(aim2, aes(fill=Category, y=Mean, x=Year)) + 
  geom_bar(position="dodge", stat="identity") +
  xlab("Follow up years") +
  ylab("Mean")
```

## Bar plots of Resource Utilization Outcomes
```{r}
prop_ed1 <- round(c(0.56603774, 0.16981132, 0.20754717, 0.05660377)*100,1)
prop_ed2 <- round(c(0.6037736, 0.1320755, NA, 0.2641509)*100,1)
prop_ed3 <- round(c(0.15094340, 0.13207547, 0.01886792, 0.69811321)*100,1)

# Combine the proportions into a data frame
cutoff = c("[0,1]", "2", "[3,5]", "NA", "[0,1]", "[2,6]",NA, "NA", "0", "1", "2", "NA")

prop_df <- data.frame(
  group=rep(c("1","2","3","4"),3),
  column = rep(c("year1", "year3", "year5"), each = 4),
  prop = c(prop_ed1, prop_ed2, prop_ed3)
)

# Create a ggplot of proportions
ggplot(prop_df, aes(x = column, y = prop, fill = group)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = paste0(prop,"%")), position = position_dodge(width = 0.9), vjust = -0.5)+
  geom_text(aes(label = cutoff,y = 0), position = position_dodge(width = 0.9))+
  labs(x = "Time", y = "Proportions", title = "Bar plots of ED visits") +
  my_theme()+
  theme(plot.title=element_text(size = 20, face = "bold"))+
  theme(legend.position = "none")
```

```{r}
prop_med1 <- round(c(0.58490566, 0.16981132, 0.18867925, 0.05660377)*100,1)
prop_med2 <- round(c(0.3962264, 0.1509434, 0.1698113, 0.2830189)*100,1)
prop_med3 <- round(c(0.16981132, 0.05660377, 0.07547170, 0.69811321)*100,1)

# Combine the proportions into a data frame
cutoff = c("[0,8]", "[9,11]", "[12,18]", "NA", "[1,5]", "[6,9]", "[10,16]", "NA", "[1,5]", "6", "[7,19]", "NA")

prop_df <- data.frame(
  group=rep(c("1","2","3","4"),3),
  column = rep(c("year1", "year3", "year5"), each = 4),
  prop = c(prop_med1, prop_med2, prop_med3)
)

# Create a ggplot of proportions
ggplot(prop_df, aes(x = column, y = prop, fill = group)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = paste0(prop,"%")), position = position_dodge(width = 0.9), vjust = -0.5)+
  geom_text(aes(label = cutoff,y = 0), position = position_dodge(width = 0.9))+
  labs(x = "Time", y = "Proportions", title = "Bar plots of medications") +
  my_theme()+
  theme(plot.title=element_text(size = 20, face = "bold"))+
  theme(legend.position = "none")
```

```{r}
prop_ad1 <- round(c(0.47169811, 0.24528302, 0.22641509, 0.05660377)*100,1)
prop_ad2 <- round(c(0.3962264, 0.1509434, 0.1886792, 0.2641509)*100,1)
prop_ad3 <- round(c(0.18867925, 0.09433962, 0.01886792, 0.69811321)*100,1)

# Combine the proportions into a data frame
cutoff = c("[0,2]", "[3,4]", "[5,9]", "NA", "[0,1]", "2", "[3,8]", "NA", "[0,1]", "2", "3", "NA")

prop_df <- data.frame(
  group=rep(c("1","2","3","4"),3),
  column = rep(c("year1", "year3", "year5"), each = 4),
  prop = c(prop_ad1, prop_ad2, prop_ad3)
)

# Create a ggplot of proportions
ggplot(prop_df, aes(x = column, y = prop, fill = group)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = paste0(prop,"%")), position = position_dodge(width = 0.9), vjust = -0.5)+
  geom_text(aes(label = cutoff,y = 0), position = position_dodge(width = 0.9))+
  labs(x = "Time", y = "Proportions", title = "Bar plots of admission") +
  my_theme()+
  theme(plot.title=element_text(size = 20, face = "bold"))+
  theme(legend.position = "none")
```

```{r}
## Comparing Outcomes by year
# knitr::kable(aim2[, c(1,3,2,4)])
```

```{r}
## Comparing Functional Status Score TOTAL by year
fun <- trach1[,21:26]

fss <- data.frame(Year=c("admission", "discharge", "year1", "year3", "year5", "mortality"), Mean=as.vector(colMeans(fun, na.rm = TRUE)))
fss$Missing <- apply(fun, 2, function(x) sum(is.na(x)))
# knitr::kable(fss)
```

## Longitudinal trajectories for FSS_total for each individual, mean (mark in red) and confidence interval
```{r}
# Reshape the data frame to long format
names(fun) <- c("admission", "discharge", "year1", "year3", "year5", "mortality")
fun1 <- cbind(x = 1:nrow(fun), fun) %>% select(-mortality)
fun_long <- tidyr::pivot_longer(fun1, -1, names_to = "group", values_to = "value")

# Extract colors from the palette
palette <- colorRampPalette(colors = c("white", "black"))
colors <- palette(265)

# Create a plot with individual curves and a mean curve
ggplot(fun_long, aes(x = factor(group, levels = c("admission", "discharge", "year1", "year3", "year5", "mortality")), y = value, group = x)) +
  my_theme() +
  theme(plot.title=element_text(size = 20, face = "bold"))+
  geom_line(aes(col = colors)) +
  # scale_color_manual(values=colors)+
  geom_point(size=1) +
  stat_summary(aes(group = 1), fun.y = "mean", geom = "line", size = 1, color = "red") +
  stat_summary(aes(color = "mean", group = "mean"), geom = "ribbon", fun = "mean", size= 1.2, alpha=0.1,
               fun.min = function(x) mean(x) - 1.96 / length(x) * sd(x), 
               fun.max = function(x) mean(x) + 1.96 / length(x) * sd(x), show.legend = FALSE) +
  labs(x = "Time", y = "FSS_total", title = "Longitudinal trajectories for FSS_total") +
  theme(legend.position = "none")

# mean((fun_long %>% filter(group=="admission"))$value)
# mean((fun_long %>% filter(group=="discharge"))$value)
# mean((fun_long %>% filter(group=="year1"))$value, na.rm=TRUE)
# mean((fun_long %>% filter(group=="year3"))$value, na.rm=TRUE)
# mean((fun_long %>% filter(group=="year5"))$value, na.rm=TRUE)

fss <- c(18.98113, 13.92453, 12.78, 10.925)
time <- c("adm", "dis", "year1", "year3")
time <- c(0,2,8,12)
cor.test(fss, time, method = "spearman")
```

## Facet plots of FSS-total 

### ED visits at 1 and 3 -year follow-up
```{r}
facet <- trach1[,c("fsstotal_year1", "fsstotal_year3", "admit_year1", "admit_year3", "edvisits_year1", "edvisits_year3", "meds_year1", "meds_year3", "ageattrach")] 

facet_ED1 <-  facet[,c(1,5,9)] %>%
  mutate(ED = ifelse(edvisits_year1==0, "ED visits = 0", ifelse(edvisits_year1==1, "ED visits = 1", "ED-visit > 1"))) %>% 
  mutate(year = "1-year follow-up") %>% 
  select(-edvisits_year1)

facet_ED1 <- facet_ED1[complete.cases(facet_ED1$ED), ]
names(facet_ED1)[names(facet_ED1) == "fsstotal_year1"] <- "fsstotal_year"


facet_ED2 <-  facet[,c(2,6,9)] %>%
  mutate(ED = ifelse(edvisits_year3==0, "ED visits = 0", ifelse(edvisits_year3==1, "ED visits = 1", "ED-visit > 1"))) %>% 
  mutate(year = "3-year follow-up") %>% 
  select(-edvisits_year3)
facet_ED2 <- facet_ED2[complete.cases(facet_ED2$ED), ]
names(facet_ED2)[names(facet_ED2) == "fsstotal_year3"] <- "fsstotal_year"

facet_ED <- rbind(facet_ED1, facet_ED2)

facet_ED <- facet_ED %>% mutate(age = ifelse(facet_ED$ageattrach<3, "age<3", ifelse(facet_ED$ageattrach>=3&facet_ED$ageattrach<8, "3<=age<8", "age>=8")))

facet_ED$ED <- factor(facet_ED$ED, levels = c("ED visits = 0", "ED visits = 1", "ED-visit > 1"))
facet_ED$age <- factor(facet_ED$age, levels = c("age<3", "3<=age<8", "age>=8"))


group.colors <- c("ED visits = 0" = "#2B83BA",  "ED visits = 1" ="#ABDDA4", "ED-visit > 1"= "#D7191C")
ed1 <- ggplot(facet_ED, aes(x=factor(year, levels = c("1-year follow-up", "3-year follow-up")), y=fsstotal_year))+ 
  my_theme() +
  theme(plot.title=element_text(size = 20, face = "bold"))+
  theme(legend.position="none") +
  geom_boxplot(aes(fill=ED))+
  labs(x = "Time", y = "FSS total") +
  facet_grid(. ~ ED) +
  scale_fill_manual(values=group.colors) +
  scale_x_discrete(guide = guide_axis(n.dodge=3))
```

### ED visits by 3 age groups: age at tracheostomy<3; 3<=age<8; age>=8
```{r}
ggplot(facet_ED, aes(x=factor(year, levels = c("1-year follow-up", "3-year follow-up")), y=fsstotal_year))+ 
  my_theme() +
  theme(legend.position="none") +
  geom_boxplot(aes(fill=ED))+
  labs(x = "Time", y = "FSS_total", title = "Facet plots of FSS-total in ED visits by 3 age groups") +
  facet_grid(age ~ ED)+
  scale_x_discrete(guide = guide_axis(n.dodge=3))
  # +geom_text(data = facet_ED%>% filter(!is.na(ED)) %>% count(ED,age), 
  #           aes(label = paste("Count:",n), y = Inf, x  = -Inf), vjust = 1, hjust = 0)
```

### Medications at 1 and 3 -year follow-up
```{r}
facet_meds1 <-  facet[,c(1,7,9)] %>%
  mutate(meds = ifelse(meds_year1<=5, "medications <= 5", ifelse(facet$meds_year1>5 & facet$meds_year1<=8, "5< medications <=8", "medications > 8"))) %>% 
  mutate(year = "1-year follow-up") %>% 
  select(-meds_year1)

facet_meds1 <- facet_meds1[complete.cases(facet_meds1$meds), ]
names(facet_meds1)[names(facet_meds1) == "fsstotal_year1"] <- "fsstotal_year"


facet_meds2 <-  facet[,c(2,8,9)] %>%
  mutate(meds = ifelse(meds_year3<=5, "medications <= 5", ifelse(facet$meds_year3>5 & facet$meds_year3<=8, "5< medications <=8", "medications > 8"))) %>% 
  mutate(year = "3-year follow-up") %>% 
  select(-meds_year3)
facet_meds2 <- facet_meds2[complete.cases(facet_meds2$meds), ]
names(facet_meds2)[names(facet_meds2) == "fsstotal_year3"] <- "fsstotal_year"

facet_meds <- rbind(facet_meds1, facet_meds2)
facet_meds <- facet_meds %>% mutate(age = ifelse(facet_meds$ageattrach<3, "age<3", ifelse(facet_meds$ageattrach>=3&facet_meds$ageattrach<8, "3<=age<8", "age>=8")))

facet_meds$meds <- factor(facet_meds$meds, levels = c("medications <= 5", "5< medications <=8", "medications > 8"))
facet_meds$age <- factor(facet_meds$age, levels = c("age<3", "3<=age<8", "age>=8"))

group.colors <- c("medications <= 5" = "#2B83BA",  "5< medications <=8" ="#ABDDA4", "medications > 8"="#D7191C")
med1 <- ggplot(facet_meds, aes(x=factor(year, levels = c("1-year follow-up", "3-year follow-up")), y=fsstotal_year))+ 
  my_theme() +
  theme(plot.title=element_text(size = 18, face = "bold"))+
  theme(legend.position="none") + 
  geom_boxplot(aes(fill=meds))+
  labs(x = "Time", y = "FSS total")+
  facet_grid(. ~ meds)+
  scale_fill_manual(values=group.colors) +
  scale_x_discrete(guide = guide_axis(n.dodge=3))
```

### Medications by 3 age groups
```{r}
ggplot(facet_meds, aes(x=factor(year, levels = c("1-year follow-up", "3-year follow-up")), y=fsstotal_year))+ 
  my_theme() +
  theme(plot.title=element_text(size = 13, face = "bold"))+
  theme(legend.position="none") +
  geom_boxplot(aes(fill=meds))+
  labs(x = "Time", y = "FSS_total", title = "Facet plots of FSS-total in medications by 3 age groups") +
  facet_grid(age ~ meds)+
  scale_x_discrete(guide = guide_axis(n.dodge=3))
```

### Hospital admissions at 1 and 3 -year follow-up
```{r}
facet_admit1 <-  facet[,c(1,3,9)] %>%
  mutate(admit = ifelse(admit_year1==0, "admissions = 0", ifelse(admit_year1==1, "admissions = 1", "admissions > 1"))) %>% 
  mutate(year = "1-year follow-up") %>% 
  select(-admit_year1)

facet_admit1 <- facet_admit1[complete.cases(facet_admit1$admit), ]
names(facet_admit1)[names(facet_admit1) == "fsstotal_year1"] <- "fsstotal_year"


facet_admit2 <-  facet[,c(2,4,9)] %>%
  mutate(admit = ifelse(admit_year3==0, "admissions = 0", ifelse(admit_year3==1, "admissions = 1", "admissions > 1"))) %>% 
  mutate(year = "3-year follow-up") %>% 
  select(-admit_year3)
facet_admit2 <- facet_admit2[complete.cases(facet_admit2$admit), ]
names(facet_admit2)[names(facet_admit2) == "fsstotal_year3"] <- "fsstotal_year"

facet_admit <- rbind(facet_admit1, facet_admit2)
facet_admit <- facet_admit %>% mutate(age = ifelse(facet_admit$ageattrach<3, "age<3", ifelse(facet_admit$ageattrach>=3&facet_admit$ageattrach<8, "3<=age<8", "age>=8")))

facet_admit$age <- factor(facet_admit$age, levels = c("age<3", "3<=age<8", "age>=8"))

group.colors <- c("admissions = 0" = "#2B83BA",  "admissions = 1" ="#ABDDA4", "admissions > 1"="#D7191C")
admit1 <- ggplot(facet_admit, aes(x=factor(year, levels = c("1-year follow-up", "3-year follow-up")), y=fsstotal_year))+ 
  my_theme() +
  theme(plot.title=element_text(size = 16, face = "bold"))+
  theme(legend.position="none") + 
  geom_boxplot(aes(fill=admit))+
  labs(x = "Time", y = "FSS total")+
  facet_grid(. ~ admit)+
  scale_fill_manual(values=group.colors) +
  scale_x_discrete(guide = guide_axis(n.dodge=3))
```

### Hospital admissions by 3 age groups
```{r}
ggplot(facet_admit, aes(x=factor(year, levels = c("1-year follow-up", "3-year follow-up")), y=fsstotal_year))+ 
  my_theme() +
  theme(plot.title=element_text(size = 11, face = "bold"))+
  theme(legend.position="none") +
  geom_boxplot(aes(fill=admit))+
  labs(x = "Time", y = "FSS_total", title = "Facet plots of FSS-total in hospital admissions by 3 age groups") +
  facet_grid(age ~ admit)+
  scale_x_discrete(guide = guide_axis(n.dodge=3))
# +
#  geom_text(data = facet_admit%>% filter(!is.na(admit)) %>% count(admit,age), 
#              aes(label = paste("Count:",n), y = Inf, x  = -Inf), vjust = 1, hjust = 0)
```

```{r}
ggarrange(admit1, ed1, med1,
                    labels = c("admissions", "ED visits", "medications"),
                    ncol = 2, nrow = 2)

ggarrange(admit1, ggarrange(ed1), nrow = 2) 

options(repr.plot.width = 5, repr.plot.height =20)

png(filename="bench_query_sort.png", width=500, height=600)
ggarrange(admit1, ed1, med1, nrow = 3)

library(patchwork)
library(gridExtra)

# Combine the plots with gridExtra
grid.arrange(admit1, ed1, med1)
```

## Summary of age, mortality, race, insurance, indication_fortrach, hospital_lengthofstay, sedatives, and vasoactives
```{r}
my.render.cont <- function(x) {
    with(stats.apply.rounding(stats.default(x), digits=2), c("",
        "Mean (SD)"=sprintf("%s (&plusmn; %s)", MEAN, SD)))
}

# table1(~ ., data=trach1[,-1], render.continuous=my.render.cont)
table1(~ ., data=trach1[,c(2, 7, 3, 8, 4, 9:14)])
```

## Summary of pcpc and fss_total
```{r}
pcpc <- trach1[,c(15,17:19)]
pcpc_long <- gather(pcpc, key = "year", value = "pcpc")
pcpc_long$year=rep(c("admit", "year1", "year3", "year5"), each=53)

fss_total <- trach1[,c(21, 23:25)]
fss_total_long <- gather(fss_total, key = "year", value = "fss_total")
data <- data.frame(year=pcpc_long$year,pcpc=pcpc_long$pcpc, fss_total=fss_total_long$fss_total)
data$pcpc <- factor(data$pcpc, levels = c("Normal", "Mild", "Moderate", "Severe", "Coma"))
table1(~ pcpc+fss_total | year, data = data, by = "year", overall=FALSE)
```

## Summary of admit, edvisits and meds
```{r}
admit <- trach1 %>% select(starts_with("admit_year"))
admit_long <- gather(admit, key = "year", value = "admit")
admit_long$year <- substr(admit_long$year, nchar(admit_long$year) - 4, nchar(admit_long$year))

edvisits <- trach1 %>% select(starts_with("edvisits"))
edvisits_long <- gather(edvisits, key = "year", value = "edvisits")

meds <- trach1 %>% select(starts_with("meds_year"))
meds_long <- gather(meds, key = "year", value = "meds")

data <- data.frame(year=admit_long$year,admit=admit_long$admit, edvisits=edvisits_long$edvisits, meds=meds_long$meds)
table1(~ admit+edvisits+meds | year, data = data, by = "year", overall=FALSE)
```
```{r}
# # quantile(trach1$admit_year1, na.rm = TRUE)
# # 0%  25%  50%  75% 100%
proptab <- function(x){
  prop.table(table(cut(x,quantile(x, na.rm = TRUE)[c(1,3,4,5)], include.lowest = TRUE, right = TRUE), exclude=NULL))
}
proptab(trach1$admit_year1)
proptab(trach1$admit_year3)
proptab(trach1$admit_year5)
proptab(trach1$edvisits_year1)
# proptab(trach1$edvisits_year3)
proptab(trach1$edvisits_year5)
proptab(trach1$meds_year1)
proptab(trach1$meds_year3)
proptab(trach1$meds_year5)


prop.table(table(cut(trach1$edvisits_year3,quantile(trach1$edvisits_year3, na.rm = TRUE)[c(1,4,5)], include.lowest = TRUE, right = TRUE), exclude=NULL))
```